cardinalR: Generating interesting high-dimensional data structures

### 2025-10-05 paper-cardinalR.Rmd
Updated branching section

The simplest structures are approximately linear branches, generated by the `gen_linearbranches(n, p, k)` function (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch) b"} else { "Figure \\@ref(fig:branch-proj) b"})`). These consist of $k$ short line segments in the first two dimensions, with added jitter to simulate variability. Mathematically, each branch $i$ is defined as

@@

To introduce curvature, the `gen_curvybranches(n, p, k)` function generates $k$ curvilinear branches in $p\text{-}D$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch) c"} else { "Figure \\@ref(fig:branch-proj) c"})`). Branches $1$ and $2$ are simple parabolas defined as

@@

An extension to the regular grid of points is to consider the points being uniformly distributed along each axis, as opposed to evenly spaced. The function `gen_unifcube(n, p)` is identical to the regular grid of points, except instead of points being placed in integer grid coordinates, they are placed at a uniformly distributed point inside the $p\text{-}D$ cube (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:cube) b"} else { "Figure \\@ref(fig:cube-proj) b"})`).

@@

The S-curve structure represents a smooth, non-linear manifold in 3D space. Using `gen_scurve(n, p)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve) a"} else { "Figure \\@ref(fig:scurve-proj) a"})`), the geometry is defined by a latent parameter $\theta \sim U(-3\pi/2, 3\pi/2)$, which controls the curvature of the manifold. The first three coordinates form the S-shape:

>>>>

The simplest structures are approximately linear branches in \gD{}, generated by the `gen_linearbranches(n, k)` function. These consist of $k$ short line segments in the first two dimensions, with added jitter to simulate variability. Mathematically, each branch $i$ is defined as

@@

To introduce curvature, the `gen_curvybranches(n, k)` function generates $k$ curvilinear branches in \gD{}. Branches $1$ and $2$ are simple parabolas defined as

@@

An extension to the regular grid of points is to consider the points being uniformly distributed along each axis, as opposed to evenly spaced. The function `gen_unifcube(n, p)` is identical to the regular grid of points, except instead of points being placed in integer grid coordinates, they are placed at a uniformly distributed point inside the $p\text{-}D$ cube.

@@

The S-curve structure represents a smooth, non-linear manifold in 3D space. Using `gen_scurve(n, p)`, the geometry is defined by a latent parameter $\theta \sim U(-3\pi/2, 3\pi/2)$, which controls the curvature of the manifold. The first three coordinates form the S-shape:

### 2025-10-05 paper-cardinalR.tex
Updated Table 3
Updated table 10
Updated branching section

\caption{\label{tab:arg-branching-tb-pdf}The main arguments for branching shape generators.}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
\toprule
Argument & Explanation\\
\midrule
n & A numeric value representing the number of points.\\
p & A numeric value representing the number of dimensions.\\
k & A numeric value representing the number of clusters.\\
\bottomrule
\end{tabular}
\end{table}

The simplest structures are approximately linear branches, generated by the \texttt{gen\_linearbranches(n,\ p,\ k)} function (Figure \ref{fig:branch-proj} b). These consist of \(k\) short line segments in the first two dimensions, with added jitter to simulate variability. Mathematically, each branch \(i\) is defined as

@@

To introduce curvature, the \texttt{gen\_curvybranches(n,\ p,\ k)} function generates \(k\) curvilinear branches in \(p\text{-}D\) (Figure \ref{fig:branch-proj} c). Branches \(1\) and \(2\) are simple parabolas defined as

@@

An extension to the regular grid of points is to consider the points being uniformly distributed along each axis, as opposed to evenly spaced. The function \texttt{gen\_unifcube(n,\ p)} is identical to the regular grid of points, except instead of points being placed in integer grid coordinates, they are placed at a uniformly distributed point inside the \(p\text{-}D\) cube (Figure \ref{fig:cube-proj} b).

@@

The S-curve structure represents a smooth, non-linear manifold in 3D space. Using \texttt{gen\_scurve(n,\ p)} (Figure \ref{fig:scurve-proj} a), the geometry is defined by a latent parameter \(\theta \sim U(-3\pi/2, 3\pi/2)\), which controls the curvature of the manifold. The first three coordinates form the S-shape:

@@

\caption{\label{tab:noise-tb-pdf}cardinalR noise dimensions generation functions}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
\toprule
Function & Explanation\\
\midrule
gen\_noisedims & Gaussian noise dimensions with optional mean and standard deviation.\\
gen\_wavydims1 & Wavy noise dimensions based on a user-specified theta sequence with added jitter.\\
gen\_wavydims2 & Wavy noise dimensions using polynomial transformations of an existing dimension vector.\\
gen\_wavydims3 & Wavy noise dimensions using a combination of polynomial and sine transformations based on the first three dimensions of a dataset.\\
\bottomrule
\end{tabular}
\end{table}

>>>>

\caption{\label{tab:arg-branching-tb-pdf}The main arguments for branching shape generators.}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2cm}>{\raggedright\arraybackslash}p{3cm}l}
\toprule
Argument & Type & Explanation\\
\midrule
n & integer & Number of points.\\
k & integer & Number of clusters.\\
\bottomrule
\end{tabular}
\end{table}

The simplest structures are approximately linear branches in $2\text{-}D${}, generated by the \texttt{gen\_linearbranches(n,\ k)} function. These consist of \(k\) short line segments in the first two dimensions, with added jitter to simulate variability. Mathematically, each branch \(i\) is defined as

@@

To introduce curvature, the \texttt{gen\_curvybranches(n,\ k)} function generates \(k\) curvilinear branches in $2\text{-}D${}. Branches \(1\) and \(2\) are simple parabolas defined as

@@

An extension to the regular grid of points is to consider the points being uniformly distributed along each axis, as opposed to evenly spaced. The function \texttt{gen\_unifcube(n,\ p)} is identical to the regular grid of points, except instead of points being placed in integer grid coordinates, they are placed at a uniformly distributed point inside the \(p\text{-}D\) cube.

@@

The S-curve structure represents a smooth, non-linear manifold in 3D space. Using \texttt{gen\_scurve(n,\ p)}, the geometry is defined by a latent parameter \(\theta \sim U(-3\pi/2, 3\pi/2)\), which controls the curvature of the manifold. The first three coordinates form the S-shape:

@@

\caption{\label{tab:noise-tb-pdf}cardinalR noise dimensions generation functions}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2.5cm}>{\raggedright\arraybackslash}p{10.5cm}}
\toprule
Function & Explanation\\
\midrule
gen\_noisedims & Gaussian noise dimensions with optional mean and standard deviation.\\
gen\_wavydims1 & Wavy noise dimensions based on a user-specified theta sequence with added jitter.\\
gen\_wavydims2 & Wavy noise dimensions using polynomial transformations of an existing dimension vector.\\
gen\_wavydims3 & Wavy noise dimensions using a combination of polynomial and sine transformations based on the first three dimensions of a dataset.\\
\bottomrule
\end{tabular}
\end{table}
